Behavior Driven Development in Haskell

### 2012-05-25 paper.tex
brush up.

In this paper, we explore how to apply BDD to Haskell.
We brought a test method from other programming language community and
integrated it with a Haskell-original test method.
In particular, we introduced {\tt doctest}~\cite{doctest} and
integrated with QuickCheck~\cite{quickcheck} as a BDD tool.
We also make use of \hspec{} inspired by rspec~\cite{rspec}.

@@

\begin{itemize}
\item Introduced {\tt doctest} from Python community to Haskell community
and integrated it with QuickCheck
(Section \ref{sec:for-users} and Section \ref{sec:doctest})
%% \item Introduced {\tt Hspec} from Ruby community to Haskell community and
%% integrated it with QuickCheck
%% (Section \ref{sec:for-developers} and Section \ref{sec:hspec})
\item Show best current practice of behavior driven development in Haskell using {\tt doctest} and \hspec{} (Section \ref{ref:Best-current-practice})
\end{itemize}

@@

BDD tools should be treat both of them and provide functionality
to disclose examples/properties for users as user documentation.
If we can archive this, BDD tools can be used for
triple purpose: design, documentation and automatic testing.

@@

@@

\begin{itemize}
\item Haskell community does not have common syntax to describe
behaviors in Haddock.
\item Haskell community does not have a BBD tool for 
examples/properties for users.
\end{itemize}

@@

Examples and properties in Haddock markup are automatically
tested by the {\tt doctest} command. We also provided a library
to do this job. Its implementation will discussed in Section~\ref{sec:doctest}.

@@

If we have some behavior that does not fit into the user documentation, but
would rather be part of a design document, we use \hspec{}.

@@

@@

\subsection{Procedure of BBD}

@@

\noindent Note that the placeholders library provide better keyword such as {\tt notImplemented} and {\tt todo}.
Next we write examples/properties for users in Haddock markup:

@@

Note that the authors are not fundamentalist.
If functions to be defined are simple enough,
we might implement them first withtout signatures.
And their signatures can be automatically
inserted by IDE (such as Emacs/vim with {\tt ghc-mod}) later
if Haskell compilers can infer correctly.
Also, we can cut and paste examples and their results
from Haskell interpreters to their Haddock markup.

>>>>

In this paper, we explore how to apply BDD to Haskell.
We brought a test method from other programming language community and
integrated it with a Haskell-original test method.
In particular, we introduced {\tt doctest}~\cite{doctest} and
integrated with QuickCheck~\cite{quickcheck} as a BDD tool.
We also make use of \hspec{} inspired by rspec~\cite{rspec} for
BDD in Haskell.

@@

\begin{itemize}
\item Introduced {\tt doctest} from Python community to Haskell community
and integrated it with QuickCheck
(Section \ref{sec:for-users} and Section \ref{sec:doctest})
%% \item Introduced {\tt Hspec} from Ruby community to Haskell community and
%% integrated it with QuickCheck
%% (Section \ref{sec:for-developers} and Section \ref{sec:hspec})
\item Show best current practice of BDD in Haskell using {\tt doctest} and \hspec{} (Section \ref{ref:Best-current-practice})
\end{itemize}

@@

BDD tools should be treat both of them and provide functionality
to extract examples/properties for users as user documentation.
If we can archive this, BDD tools can be used for
triple purpose: design, documentation and automatic testing.

@@

To describe examples/properties for developers, we can use \hspec{}.

@@

\begin{itemize}
\item Haskell community does not have common syntax to describe
examples/properties with Haddock markup syntax.
\item Haskell community does not have a BDD tool for 
examples/properties for users.
\end{itemize}

@@

Examples and properties in comments are automatically
tested by the {\tt doctest} command. We also provided a library
to do this job. Its implementation will discussed in Section~\ref{sec:doctest}.

@@

\begin{quote}
    \emph{TBD: Behavior should be replaced with examples/properties in this subsection?}
\end{quote}

If we have examples/properties that does not fit into the user documentation, but
would rather be part of a design document, we use \hspec{}.

@@

\begin{quote}
    \emph{TBD: describe the algorithm}
\end{quote}

@@

\subsection{Procedure of BDD}

@@

\noindent Note that the {\tt placeholders} library provide better keyword such as {\tt notImplemented} and {\tt todo}.
Next we write examples/properties for users in Haddock markup:

@@

Note that the authors are not fundamentalist.
If functions to be defined are simple enough,
we might implement them first withtout signatures.
And their signatures can be automatically
inserted by IDE (such as Emacs/vim with {\tt ghc-mod}) later
if Haskell compilers can infer correctly.
Also, we can copy-and-paste examples and their results
from Haskell interpreters to their Haddock markup.
